---
title: "Homework 2"
author: Yike Zhao
output: github_document
---

This is my solution to HW2.

```{r setup}
library(tidyverse)
library(readxl)
library(dplyr)
```

## Problem 1

Read the Mr. Trashwheel dataset.

```{r}
trashwheel_df =
  read_xlsx(
    "./Data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "Mr. Trash Wheel",
    range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>%
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
```

Read precipitation data for 2018 and 2017.
```{r}
precip_2018 = 
  read_excel(
    "./Data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    skip = 1
  ) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2018) %>%
  relocate(year)

precip_2017 = 
  read_excel(
    "./Data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2017 Precipitation",
    skip = 1
  ) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2017) %>%
  relocate(year)
```

Now combine annual precipitation.

```{r}
month_df = 
  tibble(
    month = 1:12,
    month_name = month.name
  )

precip_df = 
  bind_rows(precip_2018,precip_2017)

left_join(precip_df, month_df, by = "month")
```
This dataset contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, include some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Additional data sheets include month precipitation data. 

## Problem 2

Read the NYC Transit data.

```{r}
nyc_df = 
  read_csv(
   "./Data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv"
  ) %>%
  janitor::clean_names() %>%
  select(line, station_name, station_latitude, station_longitude, route1, route2, route3, route4, route5, route6, route7, route8, route9, route10, route11, entry, vending, entrance_type, ada) %>%
  mutate(entry = ifelse(entry, yes, no))
```

This dataset contains information of NYC transit. It contains `r names(nyc_df)`. It includes `r ncol(nyc_df)` columns and `r nrow(nyc_df)` rows. I think the original dataset isn't very tidy.

```{r}
nrow(distinct(nyc_df, line, station_name))
nrow(distinct(filter(nyc_df, ada == "TRUE"), line, station_name))
nrow(filter(nyc_df, vending == "NO")) / nrow(nyc_df)
```
There are 465 distinct stations here, 84 stations are ADA compliant, 9.8% of station entrances / exits without vending allow entrance.

```{r}
nyc_route =
  mutate(
  nyc_df,
  route8 = as.character(route8),
  route9 = as.character(route9),
  route10 = as.character(route10),
  route11 = as.character(route11)
  )
nyc_df_tidy = 
  pivot_longer(
    nyc_route,
    route1:route11,
    names_to = "route_name",
    values_to = "route_number"
  )
nrow(distinct(filter(nyc_df_tidy, route_number == "A"), line, station_name))
nrow(distinct(filter(nyc_df_tidy, route_number == "A", ada == "TRUE"), line, station_name))
```
60 distinct stations serve the A train. 17 of the stations are ADA compliant. 

## Problem 3

Read the pols-month data.

```{r}
pols_month = 
  read_csv(
    "./Data/pols-month.csv"
  ) %>%
  janitor::clean_names() %>%
  separate(mon, c("year", "month", "day")) %>%
  select(-c(prez_dem, prez_gop,day))

pols_month_tidy = 
  pivot_longer(
    pols_month,
    gov_gop:rep_dem,
    names_to = "president",
    values_to = "value"
  )
```

Read the snp data.

```{r}
snp_original = 
  read_csv(
    "./Data/snp.csv"
  ) %>%
  janitor::clean_names() %>%
  separate(date, c("month", "day", "year")) %>%
  select(-day) %>%
  relocate(year, month)
```



